Left-linear rules in adhesive categories

### 2024-02-07 icalp/fusione.tex
minori

\begin{definition}
	Let $\X$ be a category and $\mathcal{A}$ a  subclass of $\mor(\X)$. We say that  $\mathcal{A}$ is:
	\begin{itemize}
		\parbox{11cm}{\item 
		\emph{stable under pushouts (pullbacks)} if for every pushout (pullbacks) square as the one on the right, 	if $m \in \mathcal{A}$ ($n\in \mathcal{A}$) then $n \in \mathcal{A}$ ($m \in \mathcal{A}$);
			\item \emph{closed under composition} if $g, f\in \mathcal{A}$ implies $g\circ f\in \mathcal{A}$ whenever $g$ and $f$ are composable.}\hfill
		\parbox{2cm}{$\xymatrix{A\ar[r]^f  \ar[d]_{m}& B \ar[d]^n \\ C \ar[r]_g & D}$}
		\parbox{11cm}{}\hfill 
	\end{itemize}
\end{definition}

@@

\begin{definition}[\cite{habel2012mathcal,heindel2009category}]
	Let $\X$ be a $\mathcal{M}$-adhesive category, a  \emph{left $\mathcal{M}$-linear} rule $\rho$ is a pair $(l,r)$ of arrows with the same domain, such that $l$ belongs to $\mathcal{M}$.  The rule $\rho$ is said to be \emph{$\mathcal{M}$-linear} if $r\in \mathcal{M}$ too. A rule $\rho$ is said to be \emph{consuming} if $l$ is not an isomorphism. We will represent a rule $\rho$ as a span 
	\[\xymatrix{L & K\ar@{>->}[l]_{l} \ar[r]^{r} & R}\]
	$L$ is the \emph{left-hand side}, $R$ is the \emph{right-hand side} and $K$ the \emph{glueing object}.

@@

	Given  two objects $G$ and $H$ and a rule $\rho=(l,r)$ in $\R$, a \emph{direct derivation $\mathscr{D}$ from $G$ to $H$ applying the rule $\rho$}, is a diagram as the one below, in which both squares are pushouts. 
	\[\xymatrix{L \ar[d]_{m}& K \ar[d]^{k}\ar@{>->}[l]_{l} \ar[r]^{r} & R \ar[d]^{h}\\G & \ar@{>->}[l]^{f} D \ar[r]_{g}& H}\]
	The arrow $n$ is called the \emph{match} of the derivation, while $h$ is its \emph{back-match}.
	We will denote a direct derivation $\dder{D}$ between $G$ and $H$ as $\dder{D}\colon G\Mapsto H$. 
\end{definition}

@@

\begin{definition}
	Let $(\X, \R)$ be a left-linear DPO rewrityng system, with $\X$ $\mathcal{M}$-adhesive. The \emph{DPO-derivation graph} of $(\X, \R)$ is the (large)  directed graph $\gpo$ having as vertices the objects of $\X$ and in which an edge between $G$ and $H$ is a direct derivation $\dder{D}\colon G\Mapsto H$.	A \emph{derivation} $\der{D}$ between two objects $G$ and $H$ is a path between them in $\gpo$. The \emph{source} and \emph{target} of $\der{D}$ are, respectively, $G$ and $H$.
\end{definition}
}
\begin{remark}
We can spell out more explicitly what  a derivation $\dder{D}$ is.  An \emph{empty derivation} starting and ending in $G$ is just $G$ itself.  A \emph{non-empty derivation} $\dder{D}$ is a sequence $\{\dder{D}_i\}_{i=0}^n$ of direct derivations such that:
\begin{enumerate}
	\item for every index $i$, $\dder{D}_i$ is a direct derivation $G_i \Mapsto G_{i+1}$;
	\item $G_0=G$ and $G_{n+1}=H$.
\end{enumerate}

@@

Moreover,  if every $\dder{D}_i$ applies the rule $\rho_i\in R$, then we can define an associated sequence of rules as $r(\der{D})$ as $\{\rho_i\}_{i=0}^n$.
\end{remark}

@@

\begin{notation}
  Let $\der{D}=\{\dder{D}_i\}_{i=0}^n$ be a derivation. We will depict
  the $i^\text{th}$ element $\dder{D}_i$ of $\der{D}$ as in the
  following diagram.
  \[\xymatrix{L_i \ar[d]_{m_i}& K_i \ar[d]^{k_i}\ar@{>->}[l]_{l_i} \ar[r]^{r_i} & R_i \ar[d]^{h_i} \\G_{i} & \ar@{>->}[l]^{f_{i}} D_{i} \ar[r]_{g_{i}}& G_{i+1} }\]
  Notice that, in particular, if $\der{D}\colon G\to H$, then $G_0=G$
  and $G_{n+1}=H$. When $\der{D}$ has length $1$ we will suppress the
  indexes. In such case, we will also identify $\der{D}$ with its only
  element.
\end{notation}

@@

\rem{Non introdurrei una classe, come detto sopra}{
\begin{definition}[Tame rewriting systems]A left-linear DPO rewriting system is \emph{tame} if every independence pair between two direct derivations is good.
\end{definition}
\begin{example}\label{ex:diff2}In light of \Cref{thm:church},  \Cref{ex:diff1} provides an example of an independence pair which is not good, and thus of a non tame rewriting system.
\end{example}
\begin{remark}
	By \Cref{thm:church} the existence of a good independence pair entails switchability which, in turn, entails weak sequential independence by construction. Tame rewriting systems are exactly those rewriting systems in which these three notions coincide.
\end{remark}
}

\rem{Questo e' quanto gia' asserito sopra, mettere in appendice e riferire}{A source of tame left-linear DPO rewriting systems is given by the linear ones, as shown by the following proposition.

@@

Since $(\X, \R)$ is linear, then $r_1\colon K_1\to R_1$ belongs to $\mathcal{M}$, thus it admits a pushout along $u_1\colon K_1\to P$, as desired. Moreover, let us consider the two rectangles below.
			\[\xymatrix{K_0 \ar@/^.4cm/[rr]^{k_0}\ar@{>->}[d]_{r_0}\ar[r]_{u_0} &P\ar@{>->}[d]^{p_1} \ar@{>->}[r]_{p_0} & D_0 \ar@{>->}[d]^{f_0}& K_1 \ar@/^.4cm/[rr]^{k_1}\ar@{>->}[d]_{l_1}\ar[r]_{u_1} &P\ar@{>->}[d]^{p_0} \ar@{>->}[r]_{p_1} & D_1 \ar@{>->}[d]^{f_1}\\ 	R_0 \ar@/_.4cm/[rr]_{h_0} \ar[r]^{i_0}&D_1 \ar@{>->}[r]^{f_1} & G_1& L_1 \ar@/_.4cm/[rr]_{m_1} \ar[r]^{i_1}&D_0 \ar@{>->}[r]^{g_0} & G_1} \]
	By hypothesis $r_0$ and $l_1$ are in $\mathcal{M}$, thus $f_1$ and $g_0$ belong to it too. The first point of \Cref{lem:popb}  yields the thesis. 
	\end{proofEnd}
\end{proof}
}

@@

\appendix
\section{Properties of $\mathcal{M}$-adhesive categories}\label{app:ade}
\subsection{Some results on $\mathcal{A}$-stable and $\mathcal{A}$-Van Kampen squares}
We will start proving some general results regarding $\mathcal{A}$-Van Kampen and $\mathcal{A}$-stable squares, which, wll be used in the proof of the results of \Cref{sec:ade}. Let us begin recalling some classical results about  about pullbacks and pushouts

@@

\begin{theoremEnd}[category=sec2]{proposition}\label{prop:deco}Let  $\mathcal{A}$ be a class of arrows stable under pullbacks. For every arrow $f\colon X\to Y$ and monomorphism $m\colon Y\to Z$, if $m\circ f \in\mathcal{A}$ then $f\in \mathcal{A}$.
\end{theoremEnd}
\begin{proof}\begin{proofEnd}
		Take the diagram
		\[\xymatrix{X \ar[d]_{\id{X}}\ar[r]^{f}& Y \ar[r]^{\id{Y}}  \ar[d]_{\id{Y}}& Y \ar[d]^{m}\\
			X \ar[r]_{f}& Y \ar[r]_{m} & Z}\]
		Since $m$ is mono the right square is a pullback, while the left square is a pullback by construction. By \Cref{lem:pb1} the whole rectangle is a pullback and the thesis follows.
	\end{proofEnd}
\end{proof}
\begin{corollary}
	\label{cor:deco}
	In every $\mathcal{M}$-adhesive category $\X$, the class $\mathcal{M}$ is closed under decomposition.
\end{corollary}

@@

From \Cref{prop:pbpoad}, in turn, we can derive the following corollaries.
\begin{theoremEnd}[category=sec2]{corollary}\label{cor:rego}
	In a $\mathcal{M}$-adhesive category $\X$, every $m\in\mathcal{M}$ is a regular mono.
\end{theoremEnd}
\begin{proof}\begin{proofEnd}
		Let $m$ be an element of $\mathcal{M}$ and consider its pushout along itself.
		\[\xymatrix{X\ar@{>->}[r]^m \ar@{>->}[d]_m& Y\ar@{>->}[d]^{f}\\Y \ar@{>->}[r]_g & Z}\]
		By \Cref{prop:pbpoad} this square is a pullback, proving that $m$ is the equalizer of the arrows $f,g\colon Y\rightrightarrows Z$. 
	\end{proofEnd} 
\end{proof}

@@

\noindent
\parbox{9cm}{
	\begin{theoremEnd}[category=sec2]{lemma}\label{lem:radj}
		Let $f\colon X \to Y$ be an arrow in an $\mathcal{M}$-adhesive category $\X$ and suppose that the left square aside is a pushout while the  right one is a pullback, with $ m\colon M \rightarrowtail X$ and $n\colon N \rightarrowtail Y$ in $\mathcal{M}$.
		Then $n\leq k$ if and only if $p_2\leq m$.
\end{theoremEnd}}
\parbox{3cm}{\vspace{-0ex}
	$\xymatrix{M \ar@{>->}[d]_{m}\ar[r]^{h}& Q \ar@{>->}[d]^{k} & P \ar[r]^{p_1} \ar@{>->}[d]_{p_2}& N \ar@{>->}[d]^{n}\\X\ar[r]_{f}&Y &X \ar[r]_f&Y}$}

\begin{proof} 
	\begin{proofEnd}  
		We start pulling back $n$ along $k$ and then pulling back the resulting arrow along $h$, getting the following two pullbacks square:
		\[\xymatrix{B \ar@{>->}[d]_{b}\ar@{>->}[r]^{s}& N \ar@{>->}[d]^{n} & A \ar[r]^{r} \ar@{>->}[d]_{a}& B \ar@{>->}[d]^{b}\\Q\ar@{>->}[r]_{k}&Y &M \ar[r]_h&Q\\}\]
		Notice that 
		\[
		f\circ m\circ a=k\circ h\circ a=k\circ b\circ r=n\circ s\circ r\]

@@

		The thesis now follows immediately from \Cref{lem:varie}. \qedhere 
	\end{proofEnd}
\end{proof}

\noindent
\parbox{10cm}{ 
	\begin{theoremEnd}[category=sec2]{corollary}[Uniqueness of pushouts complements]\label{lem:pocomp} 
		Let $\X$ be a $\mathcal{M}$-adhesive category. Given $m\colon X\mto Y$ in $\mathcal{M}$ and $f\colon Y\to Z$, let $(h_1, k_1)$ and $(h_2, k_2)$ be the pushout complements of $m$  and $f$ depicted aside. Then there exists a unique isomorphism $\phi\colon W_1\to W_2$ making the diagram on the right commutative.
\end{theoremEnd}}
\parbox{3cm}{$\xymatrix{&X \ar@{>->}[r]^{m} \ar[d]_{h_1} \ar@/_.3cm/[dl]_{h_2}& Y \ar[d]^{f} \\ W_2 \ar@{>->}@/_.4cm/[rr]_{k_2} &W_1 \ar@{.>}[l]_{\phi} \ar@{>->}[r]^{k_1} & Z }$}

\begin{proof}\begin{proofEnd}
		Consider the two pushout squares
		\[\xymatrix{M \ar[r]^{h_1}\ar@{>->}[d]_{m}& W_1 \ar@{>->}[d]^{k_1} & M \ar@{>->}[d]_{m} \ar[r]^{h_1}& W_2 \ar@{>->}[d]^{k_2}\\Y\ar[r]_{f}&Z &Y \ar[r]_{f}&Z}\]
		Since $\mathcal{M}$ is in $\mathcal{M}$, \cref{prop:pbpoad} guarantees that they are both pullbacks.
		Since $m\leq m$, by \Cref{lem:radj} we get that now entails that $k_1\leq k_2$ and $k_2\leq k_1$. Thus there exists an isomorphism $\phi\colon W_1\to W_2$ such that $k_1=k_2\circ \phi$. To see that $h_2=\phi\circ h_1$, we can compute:
		\[
		k_2\circ \phi \circ h_1  = k_1\circ h_1= n\circ m= k_2\circ h_2\]

		The claim now follows since $k_2$ is a monomorphism.
	\end{proofEnd} 
\end{proof}

@@

\begin{definition}Let $\mathcal{M}$ be a class of monos in a category $\X$, closed under composition, containing all isomorphisms and stable under pullbacks and pushouts. Suppose that a  diagram
	\[\xymatrix{X \ar[d]_{a} \ar[r]^{f}& \ar[r]^{g} Y \ar[d]^{b}& Z \ar[d]^{c}\\ A \ar[r]_{h}& B \ar[r]_{k}& C}\]
	is given and the whole rectangle is a pushout. We say that $\X$ satisfies:
	\begin{itemize}
		\item the \emph{$\mathcal{M}$-mixed decomposition} property if, whenever $k$ belongs to $\mathcal{M}$ and the right half of the previous diagram is a pullback, then the left one is a pushout.
		\item the \emph{$\mathcal{M}$-pushout decomposition} property if whenever $a, b$ and $c$ belongs to $\mathcal{M}$ and the right half of the diagram above is a pushout, then its left half is a pushout too.
	\end{itemize}
\end{definition}

>>>>

\begin{definition}
	Let $\X$ be a category and $\mathcal{A}$ a  subclass of $\mor(\X)$. We say that  $\mathcal{A}$ is:
	\begin{itemize}
		\parbox{11cm}{\item 
		\emph{stable under pushouts (pullbacks)} if for every pushout (pullbacks) square as the one on the right, 	if $m \in \mathcal{A}$ ($n\in \mathcal{A}$) then $n \in \mathcal{A}$ ($m \in \mathcal{A}$);
			\item \emph{closed under composition} if $g, f\in \mathcal{A}$ implies $g\circ f\in \mathcal{A}$ whenever $g$ and $f$ are composable.{\tiny }}\hfill
		\parbox{2cm}{$\xymatrix{A\ar[r]^f  \ar[d]_{m}& B \ar[d]^n \\ C \ar[r]_g & D}$}
		\parbox{11cm}{}\hfill 
	\end{itemize}
\end{definition}

@@

\begin{definition}[\cite{habel2012mathcal,heindel2009category}]
	Let $\X$ be a $\mathcal{M}$-adhesive category, a  \emph{left $\mathcal{M}$-linear} rule $\rho$ is a pair $(l,r)$ of arrows with the same domain, such that $l$ belongs to $\mathcal{M}$.  The rule $\rho$ is said to be \emph{$\mathcal{M}$-linear} if $r\in \mathcal{M}$ too. A rule $\rho$ is said to be \emph{consuming} if $l$ is not an isomorphism.  	The object $L$ is the \emph{left-hand side}, $R$ is the \emph{right-hand side} and $K$ the \emph{glueing object}.

@@

	\noindent 
	\parbox{10cm}{\hspace{15pt}Given  two objects $G$ and $H$ and a rule $\rho=(l,r)$ in $\R$, a \emph{direct derivation $\mathscr{D}$ from $G$ to $H$ applying the rule $\rho$}, is a diagram as the one aside, in which both squares are pushouts. The arrow $m$ is called the \emph{match} of the derivation, while $h$ is its \emph{back-match}. 	We will denote a direct derivation $\dder{D}$ between $G$ and $H$ as $\dder{D}\colon G\Mapsto H$. }
	\parbox{3cm}{$\xymatrix{L \ar[d]_{m}& K \ar[d]^{k}\ar@{>->}[l]_{l} \ar[r]^{r} & R \ar[d]^{h}\\G & \ar@{>->}[l]^{f} D \ar[r]_{g}& H}$}
\end{definition}

@@

\begin{definition}
	Let $(\X, \R)$ be a left-linear DPO rewrityng system, with $\X$ $\mathcal{M}$-adhesive. The \emph{DPO-derivation graph} of $(\X, \R)$ is the (large)  directed graph $\gpo$ having as vertices the objects of $\X$ and in which an edge between $G$ and $H$ is a direct derivation $\dder{D}\colon G\Mapsto H$.	A \emph{derivation} $\der{D}$ between two objects $G$ and $H$ is a path between them in $\gpo$. The \emph{source} and \emph{target} of $\der{D}$ are, respectively, $G$ and $H$.
\end{definition}
}
\begin{definition}[Derivation]
	Let $(\X, \R)$ be a left-linear DPO rewrityng system, with $\X$ $\mathcal{M}$-adhesive.  An \emph{empty derivation} starting and ending in $G$ is just $G$ itself.  A \emph{non-empty derivation} $\dder{D}$ is a sequence $\{\dder{D}_i\}_{i=0}^n$ of direct derivations such that:
\begin{enumerate}
	\item for every index $i$, $\dder{D}_i$ is a direct derivation $G_i \Mapsto G_{i+1}$;
	\item $G_0=G$ and $G_{n+1}=H$.
\end{enumerate}

@@

Moreover,  if every $\dder{D}_i$ applies the rule $\rho_i\in R$, then we can define an associated sequence of rules as $r(\der{D})$ as $\{\rho_i\}_{i=0}^n$.
\end{definition}

@@

	\noindent 
\parbox{10cm}{\begin{notation}
  Let $\der{D}=\{\dder{D}_i\}_{i=0}^n$ be a derivation. We will depict
  the $i^\text{th}$ element $\dder{D}_i$ of $\der{D}$ as in the
  diagram on the right. Notice that, in particular, if $\der{D}$ is a derivation between $G$ and $H$, then $G_0=G$ and $G_{n+1}=H$. When $\der{D}$ has length $1$ we will suppress the
  indexes. In such case, we will also identify $\der{D}$ with its only
  element.
\end{notation} } \parbox{3cm}{$\xymatrix{L_i \ar[d]_{m_i}& K_i \ar[d]^{k_i}\ar@{>->}[l]_{l_i} \ar[r]^{r_i} & R_i \ar[d]^{h_i} \\G_{i} & \ar@{>->}[l]^{f_{i}} D_{i} \ar[r]_{g_{i}}& G_{i+1} }$}

@@

%\rem{Non introdurrei una classe, come detto sopra}{
\begin{definition}[Tame rewriting systems]A left-linear DPO rewriting system is \emph{tame} if every independence pair between two direct derivations is good.
\end{definition}
\begin{example}\label{ex:diff2}In light of \Cref{thm:church},  \Cref{ex:diff1} provides an example of an independence pair which is not good, and thus of a non tame rewriting system.
\end{example}
\begin{remark}
	By \Cref{thm:church} the existence of a good independence pair entails switchability which, in turn, entails weak sequential independence by construction. Tame rewriting systems are exactly those rewriting systems in which these three notions coincide.
\end{remark}
%}

%\rem{Questo e' quanto gia' asserito sopra, mettere in appendice e riferire}

A source of tame left-linear DPO rewriting systems is given by the linear ones, as shown by the following proposition.

@@

Since $(\X, \R)$ is linear, then $r_1\colon K_1\to R_1$ belongs to $\mathcal{M}$, thus it admits a pushout along $u_1\colon K_1\to P$, as desired. Moreover, let us consider the two rectangles below.
			\[\xymatrix{K_0 \ar@/^.4cm/[rr]^{k_0}\ar@{>->}[d]_{r_0}\ar[r]_{u_0} &P\ar@{>->}[d]^{p_1} \ar@{>->}[r]_{p_0} & D_0 \ar@{>->}[d]^{f_0}& K_1 \ar@/^.4cm/[rr]^{k_1}\ar@{>->}[d]_{l_1}\ar[r]_{u_1} &P\ar@{>->}[d]^{p_0} \ar@{>->}[r]_{p_1} & D_1 \ar@{>->}[d]^{f_1}\\ 	R_0 \ar@/_.4cm/[rr]_{h_0} \ar[r]^{i_0}&D_1 \ar@{>->}[r]^{f_1} & G_1& L_1 \ar@/_.4cm/[rr]_{m_1} \ar[r]^{i_1}&D_0 \ar@{>->}[r]^{g_0} & G_1} \]
	By hypothesis $r_0$ and $l_1$ are in $\mathcal{M}$, thus $f_1$ and $g_0$ belong to it too. The first point of \Cref{lem:popb}  yields the thesis. 
	\end{proofEnd}
\end{proof}

@@

\appendix
\section{Properties of $\mathcal{M}$-adhesive categories}\label{app:ade}

This first appendix is devoted to the proofs of some well-known results about $\mathcal{M}$-adhesive categories.

\subsection{Some results on $\mathcal{A}$-stable and $\mathcal{A}$-Van Kampen squares}
We will start proving some general results regarding $\mathcal{A}$-Van Kampen and $\mathcal{A}$-stable squares. Let us begin recalling some classical results about  about pullbacks and pushouts

@@

\begin{proposition}\label{prop:deco}Let  $\mathcal{A}$ be a class of arrows stable under pullbacks. For every arrow $f\colon X\to Y$ and monomorphism $m\colon Y\to Z$, if $m\circ f \in\mathcal{A}$ then $f\in \mathcal{A}$.
\end{proposition}
\begin{proof}
		Take the diagram
		\[\xymatrix{X \ar[d]_{\id{X}}\ar[r]^{f}& Y \ar[r]^{\id{Y}}  \ar[d]_{\id{Y}}& Y \ar[d]^{m}\\
			X \ar[r]_{f}& Y \ar[r]_{m} & Z}\]
		Since $m$ is mono the right square is a pullback, while the left square is a pullback by construction. By \Cref{lem:pb1} the whole rectangle is a pullback and the thesis follows.
\end{proof}
\begin{corollary}
	\label{cor:deco}
	In every $\mathcal{M}$-adhesive category $\X$, the class $\mathcal{M}$ is closed under decomposition.
\end{corollary}

@@

From \Cref{prop:pbpoad}, in turn, we can derive the following corollaries.
\begin{corollary}\label{cor:rego}
	In a $\mathcal{M}$-adhesive category $\X$, every $m\in\mathcal{M}$ is a regular mono.
\end{corollary}
\begin{proof}
		Let $m$ be an element of $\mathcal{M}$ and consider its pushout along itself.
		\[\xymatrix{X\ar@{>->}[r]^m \ar@{>->}[d]_m& Y\ar@{>->}[d]^{f}\\Y \ar@{>->}[r]_g & Z}\]
		By \Cref{prop:pbpoad} this square is a pullback, proving that $m$ is the equalizer of the arrows $f,g\colon Y\rightrightarrows Z$.  
\end{proof}

@@

\noindent
\parbox{9cm}{
	\begin{lemma}\label{lem:radj}
		Let $f\colon X \to Y$ be an arrow in an $\mathcal{M}$-adhesive category $\X$ and suppose that the left square aside is a pushout while the  right one is a pullback, with $ m\colon M \rightarrowtail X$ and $n\colon N \rightarrowtail Y$ in $\mathcal{M}$.
		Then $n\leq k$ if and only if $p_2\leq m$.
\end{lemma}}
\parbox{3cm}{\vspace{-0ex}
	$\xymatrix{M \ar@{>->}[d]_{m}\ar[r]^{h}& Q \ar@{>->}[d]^{k} & P \ar[r]^{p_1} \ar@{>->}[d]_{p_2}& N \ar@{>->}[d]^{n}\\X\ar[r]_{f}&Y &X \ar[r]_f&Y}$}

\begin{proof} 
		We start pulling back $n$ along $k$ and then pulling back the resulting arrow along $h$, getting the following two pullbacks square:
		\[\xymatrix{B \ar@{>->}[d]_{b}\ar@{>->}[r]^{s}& N \ar@{>->}[d]^{n} & A \ar[r]^{r} \ar@{>->}[d]_{a}& B \ar@{>->}[d]^{b}\\Q\ar@{>->}[r]_{k}&Y &M \ar[r]_h&Q\\}\]
		Notice that 
		\[
		f\circ m\circ a=k\circ h\circ a=k\circ b\circ r=n\circ s\circ r\]

@@

		The thesis now follows immediately from \Cref{lem:varie}. \qedhere 
\end{proof}

\noindent
\parbox{10cm}{ 
	\begin{corollary}[Uniqueness of pushouts complements]\label{lem:pocomp} 
		Let $\X$ be a $\mathcal{M}$-adhesive category. Given $m\colon X\mto Y$ in $\mathcal{M}$ and $f\colon Y\to Z$, let $(h_1, k_1)$ and $(h_2, k_2)$ be the pushout complements of $m$  and $f$ depicted aside. Then there exists a unique isomorphism $\phi\colon W_1\to W_2$ making the diagram on the right commutative.
\end{corollary}}
\parbox{3cm}{$\xymatrix{&X \ar@{>->}[r]^{m} \ar[d]_{h_1} \ar@/_.3cm/[dl]_{h_2}& Y \ar[d]^{f} \\ W_2 \ar@{>->}@/_.4cm/[rr]_{k_2} &W_1 \ar@{.>}[l]_{\phi} \ar@{>->}[r]^{k_1} & Z }$}

\begin{proof}
		Consider the two pushout squares
		\[\xymatrix{M \ar[r]^{h_1}\ar@{>->}[d]_{m}& W_1 \ar@{>->}[d]^{k_1} & M \ar@{>->}[d]_{m} \ar[r]^{h_1}& W_2 \ar@{>->}[d]^{k_2}\\Y\ar[r]_{f}&Z &Y \ar[r]_{f}&Z}\]
		Since $\mathcal{M}$ is in $\mathcal{M}$, \cref{prop:pbpoad} guarantees that they are both pullbacks.
		Since $m\leq m$, by \Cref{lem:radj} we get that now entails that $k_1\leq k_2$ and $k_2\leq k_1$. Thus there exists an isomorphism $\phi\colon W_1\to W_2$ such that $k_1=k_2\circ \phi$. To see that $h_2=\phi\circ h_1$, we can compute:
		\[
		k_2\circ \phi \circ h_1  = k_1\circ h_1= n\circ m= k_2\circ h_2\]

		The claim now follows since $k_2$ is a monomorphism. 
\end{proof}

@@

\noindent 
\parbox{10.4cm}{
\begin{definition}Let $\mathcal{M}$ be a class of monos in a category $\X$, closed under composition, containing all isomorphisms and stable under pullbacks and pushouts. Suppose that the  diagram on the right 	is given and the whole rectangle is a pushout. We say that $\X$ satisfies:
	\parbox{14cm}{\begin{itemize}
		\item the \emph{$\mathcal{M}$-mixed decomposition} property if, whenever $k$ belongs to $\mathcal{M}$ and the right half of the previous diagram is a pullback, then the left one is a pushout.
		\item the \emph{$\mathcal{M}$-pushout decomposition} property if whenever $a, b$ and $c$ belongs to $\mathcal{M}$ and the right half of the diagram above is a pushout, then its left half is a pushout too.
	\end{itemize}}
\end{definition}} \parbox{3cm}{\vspace{-2cm}$\xymatrix{X \ar[d]_{a} \ar[r]^{f}& \ar[r]^{g} Y \ar[d]^{b}& Z \ar[d]^{c}\\ A \ar[r]_{h}& B \ar[r]_{k}& C}$}
